\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgements}
\begin{center}\textbf{Acknowledgements}\end{center}

The CNRG lab at the University of Waterloo
has provided a rich and life-changing
environment for learning and
producing world-class research.
In particular, I would like to thank
the PhD students whose journeys
have overlapped with mine,
and who will soon be my coworkers at ABR:
Travis DeWolf, Daniel Rasmussen, Xuan Choo,
and Eric Hunsberger.
Not only have they provided
innumerable fruitful modeling discussions,
they have been great friends
for the past six years.
I would also like to thank
Terry Stewart, Jan Gosmann, Peter Blouw,
Jen Eliasmith, Gillian Martin,
and Julie DeWolf,
whose weighted sum of
technical and personal support
is remarkably high,
though the weights may vary.

I cannot overstate how grateful I am
to Chris Eliasmith,
who has not only
created the amazing environment
at the CNRG,
but is the best supervisor
a grad student could hope for.
He has been an incredible
role model for research,
a friend, a bandmate,
a hockey linemate,
and I look forward
to growing ABR together.

At the end of 2013,
Bernd J. Kr\"{o}ger emailed Chris
about whether Nengo could be used
for speech processing,
which I had just started
thinking about seriously.
I would like to thank Bernd
for being a truly wonderful collaborator,
providing a multitude of ideas
and important linguistic background
that have shaped the ideas in this thesis,
as well as being a kind person
who is a joy to work with.
In particular, I am grateful for
his comments on an early version of
some chapters of this thesis,
which are now much improved.
I would also like to thank Peter Birkholz
for his help with VocalTractLab,
and Jeff Orchard, Dan Brown,
Mathias Schulze, and Louis Goldstein
for serving on my thesis committee
and providing thoughtful questions,
intriguing comments, and helpful corrections.

My family (Cathy, David and Jason Bekolay)
have always been supportive
of my possibly terrible decision to go to grad school,
for which I am grateful.
Special thanks also for providing
a basement in which to finish this thesis.
Thanks also to a new member of my family,
Alexandra Irvine,
for help with aesthetics.

I would also like to give a shoutout
to the developers of the tools
that I use every day to get work done.
I have no idea where I would be now
Without Emacs, \LaTeX, Python,
NumPy, zsh, git, and Github.
Also, a shoutout to Arin, Dan,
Brian, Jake, and Amir
who provided necessary downtime
during the writing process.

Finally, I was lucky enough to
meet Emily Irvine shortly after starting my PhD.
Now, finishing up my PhD,
we have been together for four years
and married for one,
and they have been the most fulfilling
years of my life
thanks in large part to Emily.
I cannot thank her enough for
supporting me through this process,
and I am looking forward
to our life together sans thesis!

\cleardoublepage
